---
output:
  md_document:
    variant: markdown_github
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


#  `r ypages:::packageTitle('enrichplot', 'BioC')`

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
library("ypages")
```



`r badge_bioc_release("enrichplot", "green")`
`r badge_devel("guangchuangyu/enrichplot", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/enrichplot.svg)](https://www.bioconductor.org/packages/devel/bioc/html/enrichplot.html#since)

`r badge_download_bioc("enrichplot")`
`r badge_bioc_download("enrichplot", "total", "blue")`
`r badge_bioc_download("enrichplot", "month", "blue")`


[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![platform](http://www.bioconductor.org/shields/availability/devel/treeio.svg)](https://www.bioconductor.org/packages/devel/bioc/html/treeio.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/treeio.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/treeio/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/treeio/commits/master)




```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription('treeio')$Description)
```

[![Twitter](https://img.shields.io/twitter/url/http/shields.io.svg?style=social&logo=twitter)](https://twitter.com/intent/tweet?hashtags=enrichplot)


### Vignettes

+ [Visualization of Functional Enrichment Result](http://bioconductor.org/packages/devel/bioc/vignettes/enrichplot/inst/doc/enrichplot.html)


## Authors

Guangchuang YU <https://guangchuangyu.github.io>

School of Public Health, The University of Hong Kong

[![saythanks](https://img.shields.io/badge/say-thanks-ff69b4.svg)](https://saythanks.io/to/GuangchuangYu)
`r badger::badge_custom("follow me on", "微信", "green", "https://guangchuangyu.github.io/blog_images/biobabble.jpg")`
`r badger::badge_custom("打赏", "支付宝/微信", "green", "https://guangchuangyu.github.io/blog_images/pay_qrcode.png")`


## Installation

Get the released version from Bioconductor:

```r
## try http:// if https:// URLs are not supported
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
## biocLite("BiocUpgrade") ## you may need this
BiocManager::install("enrichplot")
```

Or the development version from github:

```r
## install.packages("devtools")
devtools::install_github("GuangchuangYu/enrichplot")
```


## Contributing

We welcome any contributions! By participating in this project you agree to
abide by the terms outlined in the [Contributor Code of Conduct](CONDUCT.md).

