---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


#  `r ypages:::packageTitle('enrichplot', 'BioC')`

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
library("ypages")
```



`r badge_bioc_release("enrichplot", "green")`
`r badge_devel("guangchuangyu/enrichplot", "green")`
[![Bioc](http://www.bioconductor.org/shields/years-in-bioc/enrichplot.svg)](https://www.bioconductor.org/packages/devel/bioc/html/enrichplot.html#since)

`r badge_download_bioc("enrichplot")`
`r badge_bioc_download("enrichplot", "total", "blue")`
`r badge_bioc_download("enrichplot", "month", "blue")`


[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![platform](http://www.bioconductor.org/shields/availability/devel/treeio.svg)](https://www.bioconductor.org/packages/devel/bioc/html/treeio.html#archives)
[![Build Status](http://www.bioconductor.org/shields/build/devel/bioc/treeio.svg)](https://bioconductor.org/checkResults/devel/bioc-LATEST/treeio/)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-green.svg)](https://github.com/GuangchuangYu/treeio/commits/master)


```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription('enrichplot')$Description)
```


For details, please visit <https://yulab-smu.top/clusterProfiler-book/> or
<https://yulab-smu.top/biomedical-knowledge-mining-book/>.


## :writing_hand: Authors


Guangchuang YU <https://guangchuangyu.github.io>


School of Basic Medical Sciences, Southern Medical University

[![Twitter](https://img.shields.io/twitter/url/http/shields.io.svg?style=social&logo=twitter)](https://twitter.com/intent/tweet?hashtags=enrichplot)
[![saythanks](https://img.shields.io/badge/say-thanks-ff69b4.svg)](https://saythanks.io/to/GuangchuangYu)
`r badger::badge_custom('follow me on', 'WeChat', 'green', 'https://guangchuangyu.github.io/blog_images/biobabble.jpg')`


## :arrow_double_down: Installation


Get the released version from Bioconductor:

```r
## try http:// if https:// URLs are not supported
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
## BiocManager::install("BiocUpgrade") ## you may need this
BiocManager::install("enrichplot")
```

Or the development version from github:

```r
## install.packages("devtools")
devtools::install_github("YuLab-SMU/enrichplot")
```


## :sparkling_heart: Contributing

We welcome any contributions! By participating in this project you agree to
abide by the terms outlined in the [Contributor Code of Conduct](CONDUCT.md).

