---
title: "Working with enrichment results from other tools"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
author: "Guangchuang Yu"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Working with results from other tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

# Overview

**enrichplot** visualizes any enrichment result stored as an
`enrichResult` (over-representation analysis) or `gseaResult`
(gene set enrichment analysis) object. Results produced by other
enrichment tools can be brought into this framework with two routes:

1. **Tool-specific importers**: `import_enrichr()`,
   `import_gprofiler2()`, `import_webgestalt()` and `import_fgsea()`
   map the output tables of popular tools to the canonical schema and
   return result objects directly.
2. **Generic constructors**: `as_enrichResult()` and
   `as_gseaResult()` (provided by the
   [enrichit](https://github.com/GuangchuangYu/enrichit) package and
   re-exported by **enrichplot**) convert any result table that follows
   the canonical column schema documented below.

Both routes return standard `enrichResult` / `gseaResult` objects, so
every **enrichplot** visualization (`dotplot()`, `barplot()`,
`cnetplot()`, `heatplot()`, `gseaplot()`, ...) works on them.

```{r}
library(enrichplot)
```

# The canonical ORA schema

The generic constructor `as_enrichResult()` requires a data.frame with
`ID`, `pvalue` and at least one of `geneID`, `Count` or `GeneRatio`.
All remaining canonical columns are derived when the necessary
information (query genes, universe, gene sets) is supplied.

| column | required | content |
|---|---|---|
| `ID` | yes | term identifier |
| `Description` | no | term name (defaults to `ID`) |
| `pvalue` | yes | raw p-value |
| `p.adjust` | no | adjusted p-value (computed with `pAdjustMethod` if absent) |
| `qvalue` | no | q-value (estimated if absent) |
| `geneID` | no | overlap genes, separated by `/` |
| `Count` | no | number of overlap genes |
| `GeneRatio` | no | `k/n`: overlap size / query size |
| `BgRatio` | no | `M/N`: gene set size / universe size |

The `gene` (query) and `universe` (background) vectors, and optionally
the `geneSets` list, provide the information needed to derive
`GeneRatio`, `BgRatio`, `RichFactor`, `FoldEnrichment` and `zScore`.

# enrichr

The `enrichR::enrichr()` function returns a list of result tables (one
per database) with columns `Term`, `Overlap`, `P.value`,
`Adjusted.P.value` and `Genes`:

```{r, eval=FALSE}
library(enrichR)
res <- enrichr(de_genes, databases = c("KEGG_2021_Human", "GO_Biological_Process_2023"))
```

`import_enrichr()` accepts either that list (use `db` to select the
database) or a single result table. Here we use a small table that
mimics the enrichr output:

```{r}
enrichr_table <- data.frame(
    Term = c("GO:0006915;;apoptotic process", "COVID-19"),
    Overlap = c("5/150", "3/60"),
    P.value = c(1.2e-4, 3.4e-3),
    Adjusted.P.value = c(0.011, 0.079),
    Genes = c("CASP3; BAX; BCL2; TP53; FAS", "ACE2; TMPRSS2; IL6"),
    stringsAsFactors = FALSE
)
de_genes <- c("CASP3", "BAX", "BCL2", "TP53", "FAS", "ACE2", "TMPRSS2", "IL6", "TNF", "MYC")

edo <- import_enrichr(enrichr_table, gene = de_genes, universe = paste0("Gene", 1:2000))
edo
```

The result is a standard `enrichResult` object, ready for any
visualization:

```{r}
dotplot(edo, showCategory = 10) + ggtitle("enrichr result")
```

```{r}
cnetplot(edo, showCategory = 2)
```

# g:Profiler

`gprofiler2::gost()` returns a list whose `result` element is a
data.frame:

```{r, eval=FALSE}
library(gprofiler2)
gostres <- gost(de_genes, organism = "hsapiens")
```

```{r}
gost_table <- data.frame(
    term_id = "GO:0006915",
    term_name = "apoptotic process",
    p_value = 1.2e-4,
    adjusted_p_value = 0.011,
    query_size = 10,
    intersection_size = 5L,
    term_size = 150L,
    effective_domain_size = 2000L,
    intersections = I(list(c("CASP3", "BAX", "BCL2", "TP53", "FAS"))),
    stringsAsFactors = FALSE
)
x <- import_gprofiler2(list(result = gost_table))
as.data.frame(x)[, c("ID", "GeneRatio", "BgRatio", "pvalue", "Count")]
```

If the table has no `query_size` column, pass the query genes with the
`gene` parameter for exact `GeneRatio` values.

# WebGestalt

`WebGestaltR()` returns a summary table with columns `geneSet`,
`description`, `size`, `overlap`, `rawPValue`, `adjPValue` and
`userIds`:

```{r, eval=FALSE}
library(WebGestaltR)
wg <- WebGestaltR(enrichMethod = "ORA", organism = "hsapiens",
                  enrichDatabase = "pathway_KEGG", interestGene = de_genes,
                  interestGeneType = "genesymbol", referenceGene = background,
                  referenceGeneType = "genesymbol")
```

```{r}
wg_table <- data.frame(
    geneSet = "hsa04210",
    description = "Apoptosis",
    size = 60L,
    overlap = 5L,
    rawPValue = 1.2e-4,
    adjPValue = 0.011,
    userIds = "CASP3;BAX;BCL2;TP53;FAS",
    stringsAsFactors = FALSE
)
x <- import_webgestalt(wg_table, gene = de_genes, universe = paste0("Gene", 1:2000))
dotplot(x, showCategory = 10)
```

# fgsea

`fgsea::fgsea()` takes a named statistics vector and a list of gene
sets, and returns a data.frame with `pathway`, `ES`, `NES`, `pval`,
`padj`, `size` and `leadingEdge`:

```{r, eval=FALSE}
library(fgsea)
data(geneList, package = "DOSE")
pathways <- split(gene_ids, term_ids)   # named list of gene sets
fgres <- fgsea(pathways, stats = geneList)
```

`import_fgsea()` additionally requires the ranked statistics vector
(`stats`), as most GSEA visualizations consume the ranked list.
Passing the `pathways` list used as input enables exact running-score
plots. Here we convert a small fgsea-style result table with ranked
statistics derived from the `DOSE` example data:

```{r}
data(geneList, package = "DOSE")
stats <- sort(geneList, decreasing = TRUE)
set1 <- names(stats)[1:40]
set2 <- names(stats)[200:240]

fgres <- data.frame(
    pathway = c("Set1", "Set2"),
    ES = c(0.62, -0.48),
    NES = c(2.1, -1.7),
    pval = c(1e-4, 0.01),
    padj = c(2e-4, 0.01),
    size = c(40L, 41L),
    stringsAsFactors = FALSE
)
fgres$leadingEdge <- list(
    names(stats)[1:10],
    rev(names(stats))[1:8]
)

gx <- import_fgsea(fgres, stats = stats, geneSets = list(Set1 = set1, Set2 = set2))
gx
```

```{r}
gseaplot(gx, geneSetID = "Set1")
```

# Arbitrary result tables

Tools that do not have a dedicated importer (e.g. DAVID, or in-house
scripts) can be converted with the generic constructors. Column
aliases (e.g. `PValue`, `term_id`, `padj`, `FDR`) are recognized
automatically:

```{r}
custom <- data.frame(
    term_id = "custom:001",
    name = "my in-house gene set",
    PValue = 0.001,
    FDR = 0.01,
    Genes = "CASP3, BAX, BCL2",
    stringsAsFactors = FALSE
)
x <- as_enrichResult(custom, gene = de_genes)
x@result[, c("ID", "Description", "geneID", "Count", "GeneRatio")]
```

# Session information

```{r}
sessionInfo()
```
